Stability concerns in random matrices

### 2018-03-17 random_matrix_stability.Rmd
fix abstract

**Analyses of dynamic systems using random matrices find that stability generally decreases with increasing system size. But in ecological communities, the magnitude of interaction effects will be highly dependent upon species generation time. The per capita effect of individuals from species with relatively short generation times will be lower than that of individuals from species with longer generation times. Interaction effects on are therefore expected to be non-random within species, with some species having consistently weaker or stronger effects on other species per capita. Here we show that by accounting for variation in generation time among species across three orders of magnitude, community stability is substantially increased. Further, the influence of variation in generation time becomes exponentially more important as community size increases. We suggest that variation in species generation time might be a key contributor to stability of large ecological communities.**

>>>>

**Analyses of dynamic systems using random matrices find that stability generally decreases with increasing system size. But in ecological communities, the rate of change of species density and the magnitude of interaction effects depend on species generation time. Per capita effects of individuals from species with short generation times will be lower than individuals from species with long generation times, and densities of species with shorter generation times will change more rapidly. Interaction effects are therefore expected to be correlated within species, with some species having consistently weaker or stronger effects on others per capita. Here we show that by accounting for variation in generation time among species, community stability is substantially increased. Further, the influence of variation in generation time becomes exponentially more important as community size increases. We suggest that variation in species generation time might be a key contributor to stability of large ecological communities.**
